---
title: "ZhangPeng-Project-2 part1"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```


```{r setup, include = FALSE}
library("tidyverse")
library("ggplot2")
library("stringr")
```

## 1. Introduction

This project uses a primary dataset that contains SAT School Participation and Performance informations limited to the state of Connecticut in 2012 and 2013.

This dataset contains data by school on student SAT scores relative to the SAT College and Career Readiness (CCR) Benchmark score of 1550 (critical reading, mathematics and writing sections combined) for the graduating classes of 2012 and 2013. According to research conducted by the College Board, a score of 1550 indicates that a student will have a 65 percent or greater likelihood of achieving a B- average or higher during the first year of college.

This dataset also provides an estimated SAT Participation Rate, which equals the number of SAT test-taking seniors in the school divided by the number of seniors enrolled in the school as a percentage. These participation rates are considered an estimate because the grade level of the SAT test-taker is self-reported by the student at the time of registration for the SAT while the total number of seniors enrolled in the school is reported to the Connecticut State Department of Education (CSDE) by the district. These are the best available estimates of SAT participation rates.

The Benchmark Performance reflect the results of only those who participate in the SAT. Because participation rates vary from school to school, the CSDE provides both the estimated Participation rates together with the Benchmark Performance in order to promote fair and valid comparisons across schools. 

So I would like to use a new index: Benchmark Meeting Rate(BMR) to compare Benchmark Performance among all schools and districts related to the dataset.This index equals the number of Benchmark-meeting seniors in a school divided by the number of seniors enrolled in the school as a percentage.

This report will try to analyze the distribution of Benchmark Performance and its trend among the schools in Connecticut in 2012 and 2013 , then give some improving suggestions. 

## 2. Import and tidy dataset

Dataset (SAT_School_Participation_and_Performance__2012-2013.csv) has been downloaded from the link:
https://catalog.data.gov/dataset/sat-school-participation-and-performance-2012-2013.

Import this dataset to a new tibble df. 

```{r}
df <- read_csv("D:/FTC2019FallSemester/datascience/R/project2/SAT_School_Participation_and_Performance__2012-2013.csv")#load the data set
glimpse(df)
```

Then parse dataset df, find "District Number" can march "School", and I will caculate change trend later, so I remove the columns: "District Number","Test-takers: Change%", "Participation Rate (estimate): Change%", and "Percent Meeting Benchmark: Change%";  moreover remame columns to "district","school","t_takes2012","t_takes2013","part_rate2012","part_rate2013","perc_mb2012","perc_mb2013"; finally drop off some rows with missing values; 

```{r cleaning, results=TRUE}

df <- df %>% select(-1, -6, -9, -12) %>% rename(district = "District", school = "School", t_takes2012 = "Test-takers: 2012", t_takes2013 = "Test-takers: 2013", part_rate2012 = "Participation Rate (estimate): 2012", part_rate2013 = "Participation Rate (estimate): 2013", perc_mb2012 = "Percent Meeting Benchmark: 2012", perc_mb2013 = "Percent Meeting Benchmark: 2013")
df <- df %>% filter(!(is.na(t_takes2012) | is.na(t_takes2013) | is.na(part_rate2012) | is.na(part_rate2013) | is.na(perc_mb2012) | is.na(perc_mb2013)))
df
```

## 3. Questions and findings

From the dataset,SAT Participation Rate is the number of SAT test-taking seniors in the school divided by the number of seniors enrolled in the school as a percentage; and variable Percent Meeting Benchmark only refects the rate of the number of Benchmark-meeting seniors in a school divided by the number of seniors who have taken SAT test in the school. We still do not know the information of how many senior students had met Benchmark from all seniors in a school. 

To evalue Benchmark Performance for each school, I choose a new index: Benchmark Meeting Rate(BMR), which comes from the number of Benchmark-meeting seniors in a school divided by the number of all seniors enrolled in the school.The formula for BMR of every school is below:
bmr = number of meeting Benchmark / number of total seniors = (t_takes*perc_mb) / (t_takes/part_rate) = pec_mb*part_rate

Because in the dataset, both perc_mb and part_rate use % as unit, the formula has been changed to :
bmr = pec_mb*part_rate*1e-4

Next with the index BMR, I figure out three questions to understand the Benchmark Performance among students in different schools and districts from Connecticut in 2012 and 2013.
Q1, what were BMRs for each school in 2012 and 2013 from Connecticut;
Q2, which 10 schools had the highest Benchmark Performance according to BMRs;
Q3, which district had the highest Benchmark Performance according to BMRs

### 3-1

Find BMRs for each school in 2012 and 2013 from Connecticut

```{r}
df1 <- df %>% select(1:4) %>% rename(`2012` = t_takes2012, `2013` = t_takes2013) %>% gather(3,4,key = "year", value = "t_takes")
df1
```

```{r}
df2 <- df %>% select(1,2,5,6) %>% rename(`2012` = part_rate2012, `2013` = part_rate2013) %>% gather(3,4,key = "year", value = "part_rate")
df2
```

```{r}
df3 <- df %>% select(1,2,7,8) %>% rename(`2012` = perc_mb2012, `2013` = perc_mb2013) %>% gather(3,4,key = "year", value = "perc_mb")
df3
```


```{r}
df4 <- df1 %>% full_join(df2,by = c("district","school","year")) %>% full_join(df3,by = c("district","school","year"))
df4 <- df4 %>% mutate(bmr = perc_mb*part_rate*1e-4)
df4
```

```{r}
df4 %>% 
  ggplot(aes(x = year, y = bmr, fill = year)) + 
  geom_boxplot() + labs(
    title = "2012-13 School Benchmark Performance in Connecticut",
    subtitle = "Median BMR < 30%",
    caption = "Graphic 1",
    y = "Benchmark Meeting Rate", x = "Year"
  ) + theme_bw()
```

From the graphic above , in 2012 the BMRs of Connecticut schools distributed from 0 to 80 percent, but in 2013 the rate went up obviously, a couple of schools' numbers almost touched 90 percent.

### 3-2

Find 10 schools had the highest Benchmark Performance according to BMRs;

```{r}
df5 <- df4 %>% filter(year == 2012) %>% arrange(desc(bmr)) %>% head(10)
df5
```

```{r}
df5 %>% ggplot(aes(x= bmr, y= fct_reorder(school,-bmr))) + geom_point() + labs(title = "2012 Benchmark Performance", x = "Benchmark Meeting Rate", y = "Top 10 Schools", caption = "Graphic 2") + theme_bw()
```

```{r}
df6 <- df4 %>% filter(year == 2013) %>% arrange(desc(bmr)) %>% head(10)
df6
```

```{r}
df6 %>% ggplot(aes(x= bmr, y= fct_reorder(school,-bmr))) + geom_point() + labs(title = "2013 Benchmark Performance", x = "Benchmark Meeting Rate", y = "Top 10 Schools", caption = "Graphic 3") + theme_bw()
```

Let's find top schools in both 2012 and 2013

```{r}
df7 <- df5 %>% semi_join(df6, by = "school") %>% select(2,3,7)
df8 <- df6 %>% semi_join(df5, by = "school") %>% select(2,3,7)
df9 <- full_join(df7,df8, by = c("school", "year", "bmr"))
df9 %>% group_by(school) %>% mutate(mean = mean(bmr)) %>% ggplot(aes(x = bmr, y = fct_reorder(school,-mean), colour = year, shape = year)) + geom_point() + labs(title = "2012-13 Benchmark Performance in Connecticut", x = "Benchmark Meeting Rate", y = "Top Schools", caption = "Graphic 4") + theme_bw()
```

### 3-3

Find districts had the highest Benchmark Performance according to BMRs

The top three districts in 2012

```{r}
df4 %>% filter(year == 2012) %>% group_by(district) %>% summarise(bmr = mean(bmr)) %>% arrange(desc(bmr)) %>% head(3)
```

The top three districts in 2013

```{r}
df4 %>% filter(year == 2013) %>% group_by(district) %>% summarise(bmr = mean(bmr)) %>% arrange(desc(bmr)) %>% head(3)
```

So there are the same three districts in both years, and let's find which schools in these districts

```{r}
df4 %>% select(1,2) %>% filter(district == c("Darien","New Canaan", "Wilton"))
```


## 4. Conclusion

In Connecticut, the percents of Benchmark-Meeting seniors out of all seniors in schools were low. For all schools median percent was lower than 30%(refer to Graphic 1). In other words, a huge amount of students would meet big challenge when they were studing during the first year of college. But the situation was changing positively. Mean percent of Benchmark-Meeting went up as well as more than 10 schools had kept BMR over 65%(refer to Graphic 4).

To helping more senior students in high schools prepared for future college life, we can think of more analysis on top Benchmark Performance schools. Specially top three schools. They were Darien High School,Wilton High School and New Canaan High School, from districts of Darien,Wilton and New Canaan respectively.In two years of 2012 and 2013, each school had over 75% senior students meeting Benchmark.

## 5. Code

 #load the data

df <- read_csv("D:/FTC2019FallSemester/datascience/R/project2/SAT_School_Participation_and_Performance__2012-2013.csv")

 #tidy the data

df <- df %>% select(-1, -6, -9, -12) %>% rename(district = "District", school = "School", t_takes2012 = "Test-takers: 2012", t_takes2013 = "Test-takers: 2013", part_rate2012 = "Participation Rate (estimate): 2012", part_rate2013 = "Participation Rate (estimate): 2013", perc_mb2012 = "Percent Meeting Benchmark: 2012", perc_mb2013 = "Percent Meeting Benchmark: 2013")
df <- df %>% filter(!(is.na(t_takes2012) | is.na(t_takes2013) | is.na(part_rate2012) | is.na(part_rate2013) | is.na(perc_mb2012) | is.na(perc_mb2013)))

 #add Benchmark Performance index-BMR column

df1 <- df %>% select(1:4) %>% rename(`2012` = t_takes2012, `2013` = t_takes2013) %>% gather(3,4,key = "year", value = "t_takes") 
df2 <- df %>% select(1,2,5,6) %>% rename(`2012` = part_rate2012, `2013` = part_rate2013) %>% gather(3,4,key = "year", value = "part_rate")
df3 <- df %>% select(1,2,7,8) %>% rename(`2012` = perc_mb2012, `2013` = perc_mb2013) %>% gather(3,4,key = "year", value = "perc_mb")
df4 <- df1 %>% full_join(df2,by = c("district","school","year")) %>% full_join(df3,by = c("district","school","year"))
df4 <- df4 %>% mutate(bmr = perc_mb*part_rate*1e-4)

 #visualize BMR vs. year variable

df4 %>% 
  ggplot(aes(x = year, y = bmr, fill = year)) + 
  geom_boxplot() + labs(
    title = "2012-13 School Benchmark Performance in Connecticut",
    subtitle = "Median BMR < 30%",
    caption = "Graphic 1",
    y = "Benchmark Meeting Rate", x = "Year"
  ) + theme_bw()
  
  #find and visualize 2012 top 10 Benchmark Performance schools

df5 <- df4 %>% filter(year == 2012) %>% arrange(desc(bmr)) %>% head(10)
df5 %>% ggplot(aes(x= bmr, y= fct_reorder(school,-bmr))) + geom_point() + labs(title = "2012 Benchmark Performance in Connecticut", x = "Benchmark Meeting Rate", y = "Top 10 Schools", caption = "Graphic 2") + theme_bw()

 #find and visualize 2013 top 10 Benchmark Performance schools

df6 <- df4 %>% filter(year == 2013) %>% arrange(desc(bmr)) %>% head(10)
df6 %>% ggplot(aes(x= bmr, y= fct_reorder(school,-bmr))) + geom_point() + labs(title = "2013 Benchmark Performance in Connecticut", x = "Benchmark Meeting Rate", y = "Top 10 Schools", caption = "Graphic 3") + theme_bw()

 #find and visualize top Benchmark Performance schools in both 2012 and 2013

df7 <- df5 %>% semi_join(df6, by = "school") %>% select(2,3,7)
df8 <- df6 %>% semi_join(df5, by = "school") %>% select(2,3,7)
df9 <- full_join(df7,df8, by = c("school", "year", "bmr"))
df9 %>% group_by(school) %>% mutate(mean = mean(bmr)) %>% ggplot(aes(x = bmr, y = fct_reorder(school,-mean), colour = year, shape = year)) + geom_point() + labs(title = "2012-13 Benchmark Performance in Connecticut", x = "Benchmark Meeting Rate", y = "Top Schools", caption = "Graphic 4") + theme_bw()

 #find top three Benchmark Performance districts in 2012

df4 %>% filter(year == 2012) %>% group_by(district) %>% summarise(bmr = mean(bmr)) %>% arrange(desc(bmr)) %>% head(3)

 #find top three Benchmark Performance districts in 2013

df4 %>% filter(year == 2013) %>% group_by(district) %>% summarise(bmr = mean(bmr)) %>% arrange(desc(bmr)) %>% head(3)

 #find schools in top three Benchmark Performance districts

df4 %>% select(1,2) %>% filter(district == c("Darien","New Canaan", "Wilton"))
 